Adopting agile in the large: motivations, success factors and challenges

### 2014-05-08 journal/article.tex
Coordinating work between many teams

@@

[P43] Transformation is described. Focus on “scaling up” part of adoption.

% --

* First cross team dependencies were reduced --> Problems with too much dependencies
* Traditional dependency management limited flexibility, which was demanded --> Inflexibility, work intensive, adversarial relationships
[P34]
From the onset, we sought to reduce cross-team dependencies as much as possible,
allowing each of the Scrum teams to “run” at their fastest pace.  However, as
the number of concurrent work streams increased, so did the number of
dependencies.  In particular, we found that adopting a SOA ran counter to our
efforts.
We quickly faced the prospect of implementing and integrating with a sizeable
set of centralized services that would be consumed by multiple project teams.
Within a few months, we had accumulated a significant number of integration
points, preventing us from delivering fully production-ready software until they
were resolved.
Accepting the fact that these dependencies would be part of the project, we
sought to create a lightweight and efficient process for resolving dependencies
and integration points.  We recognized that traditional approaches to dependency
management tend to place significant constraints on the sequence of work,
reducing the teams’ ability to independently prioritize their backlogs. 
Furthermore, highly contractual, document-centric approaches to defining
integration points are work-intensive, prone to error and can create adversarial
relationships between teams.  We wanted to establish an approach based on direct
communication and collaboration.
Our initial integration efforts involving five of the module teams of the Core
Application were reasonably successful. After some initial scrambling, the teams
self-organized and established a set of procedures for managing inter-module
integration.

* Worked decently with 5 teams and SoS --> Problems with global deployment
[P34]
However, our efforts to facilitate cross-team collaboration within the Core
Application benefited heavily from the fact that the five module teams were all
closely co-located, within the same organizational structure and employed an
active “Scrum of Scrums” model to ensure communication and coordination between
the teams.  As we pushed this approach out further across the program and the
globe, we began to experience significant challenges.

* Problem: Agile teams were thinking inside the team room walls
* Success: Thinking of the teams that will integrate to your work
* Failure: Thinking only inside the team walls
[P34]
We found that we had varying degrees of success in managing heavily-used
services.  In those cases where we were successful, it was invariably due to a
small group of people from both the consuming and providing teams taking
ownership of the services and proactively seeking out future integration points
and potential challenges.  Conversely, in those situations where such ownership
did not exist and people remained within their team walls, we invariably
experienced integration challenges: services consistently failing to meet the
consumer’s requirements and integration points that were poorly tested and
failed in system testing.

* Global effort problems --> Communication regressed to contractual --> Narrow communication channels were used: documents, e-mail
* Success: Individuals that saw effort in wide-channel communication, Individuals that thought outside team room walls
[P34]
As more and more integration occurred across the globe, we witnessed a tendency
to revert back to contractual forms of communication. As always, distributing
the development team significantly increases the challenge to directly
communicate and collaborate.  The natural tendency to rely on e-mail due to
differences in time zones frequently led to communication breakdowns and
unnecessary blockages – forming barriers to feedback. In some instances,
individuals broke through these barriers using instant messaging products, web
cams and adjusted working hours to facilitate as much direct communication as
possible.  At the same time other teams regressed, relying on documentation and
establishing an almost adversarial relationship with their integration partners.

* Large program problems --> Green builds hard to maintain
* Tools insufficient for large
[P34]
Given the size and distribution of our development team and the fact that the
program produced a single, integrated enterprise application, the complexity of
the build process and the level of effort to maintain it was significant.   As
the program grew in size, we recognized that the tool set with which we began
the project was no longer sufficient.  We began experiencing nearly continuous
broken builds due to conflation and timing issues on the build server.  At this
point, we began a lengthy initiative to replace the toolsets and revamp the
entire build process using AntHillPro and Subversion.  As a result, we created a
build process that effectively managed our myriad of dependencies and supported
a total build of the application in a reasonable time.

* Problems: Integration and dependency management
* Goal: Foster communication and collaboration instead of rigid processes
* Problems: Team avoiding external distractions, balance between team membership and project membership --> CI and green builds did not work
* Problems: Technical leads must work outside of team --> To coordinate
* Solution: Balance between own goals and project goals --> Think outside team room
* The team must think of others than them selves, too
[P34]
Over the lifetime of the project we encountered two key challenges due to the
scope and scale of this effort: integration and dependency management and
continuous integration. When addressing these challenges, we sought to adhere to
the basic principals of the Agile manifesto as much as possible: making the
delivery of working software our overriding goal and fostering communication and
collaboration instead of dictating rigid processes. We sought to empower
individuals and teams to make decisions and formulate their own processes, and
to take calculated risks without fear of reprisals.

However, we found that some of the qualities and practices that made the teams
work effectively in isolation actually reduced their effectiveness in the
large-scale environment.
For example: in smaller projects, we work aggressively to avoid external
distractions and encourage team members to focus exclusively on their individual
team goals as expressed in the backlog.  In large-scale projects, team members
must participate in cross-team activities and allocate time to project-wide
activities.  In particular, team technical leads were required to dedicate
significant portions of their time outside of the team room.  We found that many
of our team leads struggled with this task.  While they recognized the need for
cross-team communication, they resented context switching and the impact on
their time dedicated to the team’s sprint goals.
Furthermore, all of the teams struggled with balancing their identity as a team
member against their membership in the larger project. This was reflected most
keenly in our failure to sustain continuous integration and produce frequent
green builds.

Agile team must make trade-offs between stabilizing an application and
completing more story cards, the size of this project required teams to put
significantly more time into tracking down issues that were caused by other
teams or dependencies.   Many team members found this frustrating and
demoralizing.

>>>>

Moore and Spens [P34] describe the progression of the above mentioned
coordination problems. At first cross-team dependencies were attempted to be
reduced, but it became evident that the dependencies were an inherent part of
the project. A traditional approach to managing dependencies created excess
work, reduced independence and flexibility of teams, and created contract based
and adversarial relationships within the organization. There was some success
when up to 5 teams were coordinated using Scrum-of-Scrums, but that model could
not be scaled to a global level. The main problem was teams that were thinking
too much inside the team walls. When the new practices were scaled up the
communication channels narrowed and created communication breakdowns. Problems
with collaboration, dependencies, and integration were solved with presnce of
individuals with an proactive and open mindset towards working over team
boundaries. It was concluded that a balance is needed between completing new
stories from the team backlog and maintaining overall stability of the
application. [P34]

@@
